---
title: "Study 2"
output:
  html_document:
    df_print: paged
    toc: yes
---

This file reproduces the preprocessing and analysis steps of Study 2. The data are automatically imported from Github and necessary packages will be downloaded and installed if they are not yet available. 


```{r, message=F,warning=F}
rm(list=ls())
if (!require("pacman"))
  install.packages("pacman")
required_packages = c(
  "tidyverse",
  "afex",
  'brms',
  'BayesFactor',
  'emmeans',
  'gridExtra'
)
pacman::p_load(required_packages,
               update = F,
               character.only = T
)
```

```{r}
sessionInfo()
```


## Import data


```{r}
df = read.csv('https://raw.githubusercontent.com/mertensu/thinking-in-ratios/master/data_total_study2.csv')
```

## Preprocessing

```{r}
# replace zero ratings with 0.001
df[df$brightness_rating==0,'brightness_rating'] = 0.001
# df = df[df$brightness_rating!=0, ]


# compute correct binary choice
df = df %>% mutate(brightness_dichotom_correct = ifelse(cd < 10, 0, 1))

df$correct_response = NA
df[df$condition == 'unidirectional', 'correct_response'] = df[df$condition == 'unidirectional', 'brightness_dichotom_correct'] == df[df$condition == 'unidirectional', 'brightness_dichotom']

# remove wrong binary choice(s) in unidirectional condition
df = df %>% filter(!(correct_response == F &
                       condition == 'unidirectional'))

df[, "brightness_rating_stevens"] = ifelse((df$condition == "unidirectional") &
                                             (df$brightness_dichotom_correct == 0),
                                           100 / df$brightness_rating,
                                           df$brightness_rating
)
df[, "brightness_rating"] = ifelse((df$condition == "standard") &
                                     (df$brightness_dichotom_correct == 0),
                                   10 * (10 / df$brightness_rating),
                                   df$brightness_rating
)

df$log_brightness_rating = log(df$brightness_rating_stevens)

df$cd_factor = factor(df$cd, levels = c(1, 1.8, 3.2, 5.7, 17.9, 32.0, 57.2, 100.0))

df$brightness_rating_stevens = df$brightness_rating_stevens/100
df$cd = df$cd/100
```


## Demographics

```{r}
psych::describe(df$age)

df %>% distinct(File, .keep_all = T) %>% group_by(gender) %>% summarise(
  N = n(),
  Min =
    min(age),
  Max =
    max(age),
  Mean =
    mean(age),
  Sd =
    sd(age)
)

df %>% distinct(File, .keep_all = T) %>% count(student)
df %>% distinct(File, .keep_all = T) %>% filter(student == 1) %>% count(psycho)

df %>% distinct(File, .keep_all = T) %>% count(condition)
```

## Analysis

### ANOVA I (mixed 2(method) x 8(luminance))

#### frequentist fit

```{r}
(
  fit = aov_ez(
    dv = "log_brightness_rating",
    within = 'cd_factor',
    between = "condition",
    id = "File",
    data = df
  )
)
```

#### bayesian fit

```{r}
df$File = factor(df$File)
df$condition = factor(df$condition)
bfs = anovaBF(
  log_brightness_rating ~ cd_factor * condition + File,
  whichRandom = 'File',
  whichModels = 'top',
  data = df
)

# BF cd_factor
bf_1 = lmBF(log_brightness_rating ~ condition + File,
            whichRandom = 'File',
            data = df)
bf_2 = lmBF(
  log_brightness_rating ~ cd_factor + condition + File,
  whichRandom = 'File',
  data = df
)
(bf_cd_factor = bf_2 / bf_1)
print(paste0('logBF ', bf_cd_factor@bayesFactor$bf))

# BF condition
bf_1 = lmBF(log_brightness_rating ~ cd_factor + File,
            whichRandom = 'File',
            data = df)
bf_2 = lmBF(
  log_brightness_rating ~ cd_factor + condition + File,
  whichRandom = 'File',
  data = df
)
(bf_condition = bf_2 / bf_1)
print(paste0('logBF ', bf_condition@bayesFactor$bf))

# BF interaction
bf_1 = lmBF(
  log_brightness_rating ~ cd_factor + condition + File,
  whichRandom = 'File',
  data = df
)
bf_2 = lmBF(
  log_brightness_rating ~ cd_factor * condition + File,
  whichRandom = 'File',
  data = df
)
(bf_interaction = bf_2 / bf_1)
print(paste0('logBF ', bf_interaction@bayesFactor$bf))
```

#### Figure 2

```{r}
scaleFUN <- function(x)
  sprintf("%.1f", x)

grid = data.frame(emmeans(fit,  ~ cd_factor + condition))

ggplot(grid, aes(
  x = cd_factor,
  y = exp(emmean),
  group = condition
)) +
  geom_pointrange(aes(
    ymin = exp(lower.CL),
    ymax = exp(upper.CL),
    color = condition
  ), size =
    0.3) +
  scale_y_continuous(
    trans = 'log2',
    breaks = c(1.0, 10.0, 100.0),
    limits = c(1.0, 100.0),
    labels = scaleFUN
  ) +
  geom_line(linetype = 'dashed') +
  labs(y = 'Brightness judgement') +
  xlab(expression(paste("cd/", m ^ 2, sep = ""))) +
  scale_color_manual(
    values = c("darkgrey", "black"),
    name = "method",
    labels = c("unidirectional", "standard")
  ) +
  scale_x_discrete(labels = substring(grid$cd_factor, 2)) +
  theme_classic() +
  theme(
    legend.position = c(0.2, 0.8),
    legend.background = element_rect(color = "black")
  )
```

```{r}
ggsave(
  paste0("final_plots/study2_figure2.png"),
  dpi = 600,
  height = 4,
  width = 5,
  units = "in"
)
```



### Bayesian mixed effects models (standard condition)

#### Fit power law

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
                    set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```

```{r}
(formula = bf(brightness_rating_stevens ~ a * cd ^ b, a ~ (1 |File), b ~ (1 | File), nl = TRUE))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'standard', ], prior = prior_settings)
```


```{r}
(power_law_standard <-
  brm(formula,
    data = df[df$condition == 'standard', ],
    save_all_pars = T,
    sample_prior = 'yes',
    warmup = 2000,
    iter = 7000,
    cores = 4,
    control = list(adapt_delta = 0.96, max_treedepth = 20),
    prior = prior_settings,
    file = 'study2_power_standard_x'))
```

```{r}
(
  pow_standard_plot = pp_check(power_law_standard) +
    labs(title = 'Power law', y = 'Standard method') +
    scale_x_continuous(labels = c(0, 25, 50, 75, 100)) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = c(0.85, 0.85),
    ) 
)
```

#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(power_law_standard)
brms::waic(power_law_standard)
brms::bridge_sampler(power_law_standard)
```


#### Fit linear model

```{r}
prior_settings <-
  c(
    set_prior('normal(0,1)', class = 'b', coef = 'cd'),
    set_prior('normal(0,5)', class = 'b', coef = 'Intercept')
  )
```

```{r}
(formula = brightness_rating_stevens ~ 0 + Intercept + cd + (1 + cd | File))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'standard', ], prior = prior_settings)
```

```{r}
(linear_standard <-
  brm(
    formula,
    data = df[df$condition == 'standard', ],
    save_all_pars = T,
    sample_prior = 'yes',
    warmup = 2000,
    cores = 4,
    prior = prior_settings,
    iter = 7000,
    file = 'study2_linear_standard_x'))
```

```{r}
(
  lin_standard_plot = pp_check(linear_standard) + labs(title = 'Linear model') +
    scale_x_continuous(labels = c(0, 25, 50, 75, 100)) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = c(0.85, 0.85),
    )
)
```


#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(linear_standard)
brms::waic(linear_standard)
brms::bridge_sampler(linear_standard)
```


#### Model comparison (Bayes factors)

Compute logBF in favor of power law. Takes a few minutes to run, hence commented out.

```{r,message=F}
bfs_standard = vector('numeric', length = 5)
for (i in 1:5) {
  bfs_standard[i] = log(1 / bayes_factor(linear_standard, power_law_standard)$bf)
}
print(paste0('Mean logBF:', mean(bfs_standard), 'Std logBF:', sd(bfs_standard)))
```



### Bayesian mixed-effects models (unidirectional condition)

#### Fit power law

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
                    set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0)
                    )
```

```{r}
(formula = bf(brightness_rating_stevens ~ a * cd ^ b, a ~ (1 |File), b ~ (1 | File),  nl = TRUE))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'unidirectional', ], prior = prior_settings)
```


```{r}
(power_law_unidirectional <-
  brm(formula,
    data = df[df$condition == 'unidirectional', ],
    save_all_pars = T,
    sample_prior = 'yes',
    warmup = 2000,
    iter = 7000,
    cores = 4,
    control = list(adapt_delta = 0.96, max_treedepth = 20),
    prior = prior_settings,
    file = 'study2_power_unidirectional_x'))
```


```{r}
(
  pow_unidirectional_plot = pp_check(power_law_unidirectional) +
    labs(y = 'Unidirectional method') +
    scale_x_continuous(labels = c(0, 25, 50, 75, 100), breaks=c(0,0.25,0.5,0.75,1)) +
    theme(legend.position = c(0.85, 0.85))
) 
```



#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(power_law_unidirectional)
brms::waic(power_law_unidirectional)
brms::bridge_sampler(power_law_unidirectional)
```

#### Fit linear model

```{r}
prior_settings <-
  c(
    set_prior('normal(0,1)', class = 'b', coef = 'cd'),
    set_prior('normal(0,5)', class = 'b', coef = 'Intercept')
  )
```

```{r}
(formula = brightness_rating_stevens ~ 0 + Intercept + cd + (1 + cd | File))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'unidirectional', ], prior = prior_settings)
```

```{r}
(linear_unidirectional <-
  brm(
    formula,
    data = df[df$condition == 'unidirectional', ],
    save_all_pars = T,
    sample_prior = 'yes',
    warmup = 2000,
    cores = 4,
    prior = prior_settings,
    iter = 7000,
    file = 'study2_linear_unidirectional_x'))
```

```{r}
(
  lin_unidirectional_plot = pp_check(linear_unidirectional) +
    scale_x_continuous(labels = c(0, 25, 50, 75, 100), breaks=c(0,0.25,0.5,0.75,1)) +
    theme(legend.position = c(0.85, 0.85))
)
```

#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(linear_unidirectional)
brms::waic(linear_unidirectional)
brms::bridge_sampler(linear_unidirectional)
```

#### Model comparison (Bayes factors)

Compute logBF in favor of power law. Takes a few minutes to run, hence commented out.

```{r,message=F}
bfs_unidirectional = vector('numeric', length = 5)
for (i in 1:5) {
  bfs_unidirectional[i] = log(1 / bayes_factor(linear_unidirectional, power_law_unidirectional)$bf)
}
print(paste0('Mean logBF:', mean(bfs_unidirectional), 'Std logBF:', sd(bfs_unidirectional)))
```

#### Posterior predictive plot (Figure 3)

```{r}
pp_grid = (
  grid.arrange(
    pow_standard_plot,
    lin_standard_plot,
    pow_unidirectional_plot,
    lin_unidirectional_plot
  )
)
```

```{r}
ggsave(
  paste0("final_plots/study2_pp_plot.png"),
  plot = pp_grid,
  dpi = 600,
  height = 6,
  width = 7,
  units = "in"
)
```


### Bayesian mixed-effects models (aggregated data)

#### Fit power law (standard condition)


```{r}
df_agg = df %>% group_by(cd,condition) %>% summarise(median_rating = median(brightness_rating_stevens))

```

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
            set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```


```{r}
(formula = bf(median_rating ~ a * cd ^ b , a ~ 1, b ~ 1 , nl = TRUE))
```


```{r}
# make_stancode(formula, data = df_agg[df_agg$condition == 'standard', ], prior = prior_settings)
```


```{r}
(
  power_law_standard_agg <-
    brm(
      formula,
      data = df_agg[df_agg$condition == 'standard',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 20),
      prior = prior_settings,
      file = 'study2_power_standard_agg_x'
    )
)
```

#### Fit power law on aggregated data (unidirectional condition)


```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
            set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```


```{r}
(formula = bf(median_rating ~ a * cd ^ b , a ~ 1, b ~ 1 , nl = TRUE))
```


```{r}
# make_stancode(formula, data = df_agg[df_agg$condition == 'unidirectional', ], prior = prior_settings)
```


```{r}
(
  power_law_unidirectional_agg <-
    brm(
      formula,
      data = df_agg[df_agg$condition == 'unidirectional',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 20),
      prior = prior_settings,
      file = 'study2_power_unidirectional_agg_x'
    )
)
```


## Supplemental analysis

```{r}
df_supp = read.csv('https://raw.githubusercontent.com/mertensu/thinking-in-ratios/master/data_total_supp.csv')
```



```{r}
df_supp[df_supp$brightness_rating == 0, 'brightness_rating'] = 0.001

df_supp = df_supp %>% mutate(brightness_dichotom_correct = ifelse(cd < 10, 0, 1))


df_supp$correct_response = NA
df_supp[,'correct_response'] = df_supp['brightness_dichotom_correct'] == df_supp[, 'brightness_dichotom']

df_supp$cd_factor = factor(df_supp$cd, levels = c(1, 1.8, 3.2, 5.7, 17.9, 32.0, 57.2, 100.0))


# use same column name as in Study 2
df_supp$brightness_rating_stevens = df_supp$brightness_rating
df_supp$log_brightness_rating = log(df_supp$brightness_rating_stevens)

df_supp$brightness_rating_stevens = df_supp$brightness_rating_stevens/100
df_supp$cd = df_supp$cd/100
```



```{r}
psych::describe(df_supp$age)

df_supp %>% distinct(File, .keep_all = T) %>% group_by(gender) %>% summarise(
  N = n(),
  Min =
    min(age),
  Max =
    max(age),
  Mean =
    mean(age),
  Sd =
    sd(age)
)

df_supp %>% distinct(File, .keep_all = T) %>% count(student)
df_supp %>% distinct(File, .keep_all = T) %>% filter(student == 1) %>% count(psycho)

df_supp %>% distinct(File, .keep_all = T) %>% count(condition)
```



```{r}
df_s2 = df
df_s2$File = paste0(df_s2$File,'study2')
df_tot = rbind(df_supp,df_s2)
df_tot = df_tot %>% filter(condition != 'unidirectional')
```


```{r}
(
  fit = aov_ez(
    dv = "log_brightness_rating",
    within = 'cd_factor',
    between = "condition",
    id = "File",
    data = df_tot
  )
)
```


```{r}
df_tot$File = factor(df_tot$File)
df_tot$condition = factor(df_tot$condition)
bfs = anovaBF(log_brightness_rating ~ cd_factor * condition + File, 
              whichRandom = 'File',whichModels = 'top',data=df_tot)

# BF cd_factor
bf_1 = lmBF(log_brightness_rating ~ condition + File, whichRandom = 'File',data=df_tot)
bf_2 = lmBF(log_brightness_rating ~ cd_factor + condition + File, whichRandom = 'File',data=df_tot)
(bf_cd_factor = bf_2 / bf_1)
print(paste0('logBF ', bf_cd_factor@bayesFactor$bf))

# BF condition
bf_1 = lmBF(log_brightness_rating ~ cd_factor + File, whichRandom = 'File',data=df_tot)
bf_2 = lmBF(log_brightness_rating ~ cd_factor + condition + File, whichRandom = 'File',data=df_tot)
(bf_condition = bf_2 / bf_1)
print(paste0('logBF ', bf_condition@bayesFactor$bf))

# BF interaction
bf_1 = lmBF(log_brightness_rating ~ cd_factor + condition + File, whichRandom = 'File',data=df_tot)
bf_2 = lmBF(log_brightness_rating ~ cd_factor * condition + File, whichRandom = 'File',data=df_tot)
(bf_int = bf_2 / bf_1)
print(paste0('logBF ', bf_int@bayesFactor$bf))
```


####  Fit power law

```{r}
prior_settings <-
  c(set_prior('normal(0, 5)', nlpar = "a"),
    set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```


```{r}
(formula = bf(brightness_rating_stevens ~ a * cd ^ b , a ~ (1 |File), b ~ (1 | File), nl = TRUE))
```

```{r}
# make_stancode(formula, data = df_supp, prior = prior_settings)
```


```{r}
(
  power_law_supp <-
    brm(
      formula,
      data = df_supp,
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 20),
      prior = prior_settings,
      file = 'power_binary_standard'
    )
)
```

