---
title: "Study 3"
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
---

This file reproduces the preprocessing and analysis steps of Study 3. The data are automatically imported from Github and necessary packages will be downloaded and installed if they are not yet available. 


Create directory to save plots:

```{r}
if (!dir.exists('final_plots')) {dir.create('final_plots')}
```


```{r, echo=F,message=F,warning=F}
rm(list=ls())
if (!require("pacman"))
  install.packages("pacman")
required_packages = c(
  "tidyverse",
  "afex",
  'brms',
  'BayesFactor',
  'emmeans',
  'gridExtra'
)
pacman::p_load(required_packages,
               update = F,
               character.only = T
)
```

```{r}
set.seed(42)
```


```{r}
sessionInfo()
```


## Import data


```{r}
github_link = 'https://raw.githubusercontent.com/mertensu/thinking-in-ratios/master/'
file_name = 'data_total_study3.csv' 
df = read.csv(paste0(github_link, file_name))
```

## Preprocessing


```{r}
# replace zero ratings with 0.001
df[df$redness_rating == 0, 'redness_rating'] = 0.001


# compute correct binary choice
df = df %>% mutate(redness_correct = ifelse(saturation < 50, 0, 1))

df$correct_response = NA
df[df$condition == 'unidirectional', 'correct_response'] = df[df$condition == 'unidirectional', 'redness_correct'] == df[df$condition == 'unidirectional', 'redness_dichotom']

# remove wrong binary choice(s) in unidirectional condition
df = df %>% filter(!(correct_response == F &
                       condition == 'unidirectional'))

df[, "redness_rating_comparable"] = ifelse((df$condition == "unidirectional") &
                                             (df$redness_dichotom == 0),
                                           100 / df$redness_rating,
                                           df$redness_rating
)

df$saturation_factor = factor(df$saturation, levels = c(15, 25, 35, 45, 55, 65, 75, 85))

df$log_redness_rating = log(df$redness_rating_comparable)
# divide saturation by 100 to enable more efficient sampling
df$saturation = df$saturation/100
df$redness_rating_comparable = df$redness_rating_comparable/100

```


## Demographics

```{r}
psych::describe(df$age)

df %>% distinct(File, .keep_all = T) %>% group_by(gender) %>% summarise(
  N = n(),
  Min =
    min(age),
  Max =
    max(age),
  Mean =
    mean(age),
  Sd =
    sd(age)
)

df %>% distinct(File, .keep_all = T) %>% count(student)
df %>% distinct(File, .keep_all = T) %>% filter(student == 1) %>% count(psycho)

df %>% distinct(File, .keep_all = T) %>% count(condition)
```


## Analysis

### ANOVA I (mixed 2(method) x 8(saturation))

#### frequentist fit

```{r}

(
  fit = aov_ez(
    dv = "log_redness_rating",
    within = 'saturation_factor',
    between = "condition",
    id = "File",
    data = df
  )
)
```

#### bayesian fit

```{r}
df$File = factor(df$File)
df$condition = factor(df$condition)
(
  bfs = anovaBF(
    log_redness_rating ~ saturation_factor * condition + File,
    whichRandom = 'File',
    whichModels = 'top',
    data = df
  )
)

# BF saturation_factor
bf_1 = lmBF(log_redness_rating ~ condition + File,
            whichRandom = 'File',
            data = df)
bf_2 = lmBF(
  log_redness_rating ~ saturation_factor + condition + File,
  whichRandom = 'File',
  data = df
)
(bf_saturation_factor = bf_2 / bf_1)
print(paste0('logBF ', bf_saturation_factor@bayesFactor$bf))

# BF condition
bf_1 = lmBF(
  log_redness_rating ~ saturation_factor + File,
  whichRandom = 'File',
  data = df
)
bf_2 = lmBF(
  log_redness_rating ~ saturation_factor + condition + File,
  whichRandom = 'File',
  data = df
)
(bf_condition = bf_2 / bf_1)
print(paste0('logBF ', bf_condition@bayesFactor$bf))

# BF interaction
bf_1 = lmBF(
  log_redness_rating ~ saturation_factor + condition + File,
  whichRandom = 'File',
  data = df
)
bf_2 = lmBF(
  log_redness_rating ~ saturation_factor * condition + File,
  whichRandom = 'File',
  data = df
)
(bf_interaction = bf_2 / bf_1)
print(paste0('logBF ', bf_interaction@bayesFactor$bf))
```

#### Figure 4

```{r}
grid = data.frame(emmeans(fit,  ~ saturation_factor + condition))
grid$condition = factor(grid$condition,levels=c('unidirectional', 'standard'))

ggplot(grid, aes(
  x = saturation_factor,
  y = exp(emmean),
  group = condition
)) +
  geom_pointrange(aes(
    ymin = exp(lower.CL),
    ymax = exp(upper.CL),
    color = condition
  ), size =
    0.3) +
  scale_y_continuous(
    trans = 'log2',
    breaks = c(0.1, 1, 10, 100),
    limits = c(0.07, 100)
  ) +
  geom_line(linetype = 'dashed') +
  labs(y = 'Redness judgement', x = 'redness saturation (in %)') +
  scale_color_manual(
    values = c("darkgrey", "black"),
    name = "method",
  ) +
  scale_x_discrete(labels = substring(grid$saturation_factor, 2)) +
  theme_classic() +
  theme(
    legend.position = c(0.2, 0.85),
    legend.background = element_rect(color = "black")
  )
```
```{r}
ggsave(
  paste0("final_plots/study3_figure3.png"),
  dpi = 600,
  height = 4,
  width = 5,
  units = "in"
)
```


### Bayesian mixed effects models (standard condition)


#### Fit power law

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
                    set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```

```{r}
(formula = bf(
      redness_rating_comparable ~ a * saturation ^ b ,
      a ~ (1 | File),
      b ~ (1 | File),
      nl = TRUE
    ))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'standard', ], prior = prior_settings)
```


```{r}
(
  power_law_standard <-
    brm(
      formula,
      data = df[df$condition == 'standard',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 15),
      prior = prior_settings,
      file = 'study3_power_standard_x'
    )
)
```

```{r}
(
  pow_standard_plot = pp_check(power_law_standard) +
    labs(title = 'Power law', y = 'Standard method') +
    scale_x_continuous(labels=c(0,20,40,60), breaks=c(0,0.2,0.4,0.6)) + 
    scale_y_continuous(limits = c(0, 7)) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = c(0.9, 0.9)
    )
)
```

#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(power_law_standard)
brms::waic(power_law_standard)
brms::bridge_sampler(power_law_standard)
```


#### Fit linear model


```{r}
prior_settings = c(
  set_prior('normal(0,1)', class = 'b', coef = 'saturation'),
  set_prior('normal(0,5)', class = 'b', coef = 'Intercept')
)
```

```{r}
(formula = redness_rating_comparable ~ 0 + Intercept + saturation + (1 + saturation |File))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'standard', ], prior = prior_settings)
```

```{r}
(linear_standard <-
  brm(
    formula,
    data = df[df$condition == 'standard', ],
    save_all_pars = T,
    sample_prior = 'yes',
    warmup = 2000,
    cores = 4,
    prior = prior_settings,
    iter = 7000,
    file = 'study3_linear_standard_x'))
```

```{r}
(
  lin_standard_plot = pp_check(linear_standard) + labs(title = 'Linear model') + 
    scale_x_continuous(labels=c(0,20,40,60), breaks=c(0,0.2,0.4,0.6)) + 
    scale_y_continuous(limits = c(0, 7)) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = c(0.9, 0.9)
  )
)
```

#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(linear_standard)
brms::waic(linear_standard)
brms::bridge_sampler(linear_standard)
```


#### Model comparison (Bayes factors)

Compute logBF in favor of power law. Takes a few minutes to run, hence commented out.

```{r,message=F}
bfs_standard = vector('numeric', length = 5)
for (i in 1:5) {
  bfs_standard[i] = log(1 / bayes_factor(linear_standard, power_law_standard)$bf)
}
print(paste0('Mean logBF:', mean(bfs_standard), 'Std logBF:', sd(bfs_standard)))
```

### Bayesian mixed effects models (unidirectional condition)


#### Fit power law

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
                    set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```

```{r}
(formula = bf(
      redness_rating_comparable ~ a * saturation ^ b ,
      a ~ (1 | File),
      b ~ (1 | File),
      nl = TRUE
    ))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'unidirectional', ], prior = prior_settings)
```


```{r}
(
  power_law_unidirectional <-
    brm(
      formula,
      data = df[df$condition == 'unidirectional',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 15),
      prior = prior_settings,
      file = 'study3_power_unidirectional_x'
    )
)
```

```{r}
(
  pow_unidirectional_plot = pp_check(power_law_unidirectional) +
    scale_x_continuous(labels=c(0,20,40,60), breaks=c(0,0.2,0.4,0.6)) + 
    scale_y_continuous(limits = c(0, 7)) +
    labs(y = 'Unidirectional method') +
    theme(legend.position = c(0.9, 0.9))
) 
```



#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(power_law_unidirectional)
brms::waic(power_law_unidirectional)
brms::bridge_sampler(power_law_unidirectional)
```

#### Fit linear model

```{r}
prior_settings = c(
  set_prior('normal(0,1)', class = 'b', coef = 'saturation'),
  set_prior('normal(0,5)', class = 'b', coef = 'Intercept')
)
```

```{r}
(formula = redness_rating_comparable ~ 0 + Intercept + saturation + (1 + saturation |File))
```

```{r}
# make_stancode(formula, data = df[df$condition == 'unidirectional', ], prior = prior_settings)
```

```{r}
(
  linear_unidirectional <-
    brm(
      formula,
      data = df[df$condition == 'unidirectional',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      cores = 4,
      prior = prior_settings,
      iter = 7000,
      file = 'study3_linear_unidirectional_x'
    )
)
```

```{r}
(lin_unidirectional_plot = pp_check(linear_unidirectional) +
   scale_x_continuous(labels=c(0,20,40,60), breaks=c(0,0.2,0.4,0.6)) + 
   scale_y_continuous(limits = c(0, 7)) +
   theme(legend.position = c(0.9, 0.9)))
```

#### Model metrics (LOOCV, WAIC, marginal likelihood)

```{r}
brms::loo(linear_unidirectional)
brms::waic(linear_unidirectional)
brms::bridge_sampler(linear_unidirectional)
```


#### Model comparison (Bayes factors)

Compute logBF in favor of power law. Takes a few minutes to run, hence commented out.

```{r,message=F}
bfs_unidirectional = vector('numeric', length = 5)
for (i in 1:5) {
  bfs_unidirectional[i] = log(1 / bayes_factor(linear_unidirectional, power_law_unidirectional)$bf)
}
print(paste0('Mean logBF:', mean(bfs_unidirectional), 'Std logBF:', sd(bfs_unidirectional)))
```



#### Posterior predictive plot (Figure 5)

```{r}
pp_grid = (
  grid.arrange(
    pow_standard_plot,
    lin_standard_plot,
    pow_unidirectional_plot,
    lin_unidirectional_plot
  )
)
```


```{r}
ggsave(
  paste0("final_plots/study3_pp_plot.png"),
  plot = pp_grid,
  dpi = 600,
  height = 6,
  width = 7,
  units = "in"
)
```



### Bayesian mixed-effects models (aggregated data)

#### Fit power law (standard condition)


```{r}
df_agg = df %>% group_by(saturation,condition) %>% summarise(median_rating = median(redness_rating_comparable))

```

```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
            set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```


```{r}
(formula = bf(median_rating ~ a * saturation ^ b , a ~ 1, b ~ 1 , nl = TRUE))
```


```{r}
# make_stancode(formula, data = df_agg[df_agg$condition == 'standard', ], prior = prior_settings)

```


```{r}
(
  power_law_standard_agg <-
    brm(
      formula,
      data = df_agg[df_agg$condition == 'standard',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 20),
      prior = prior_settings,
      file = 'study3_power_standard_agg_x'
    )
)
```

#### Fit power law (unidirectional condition)



```{r}
prior_settings <- c(set_prior('normal(0, 5)', nlpar = "a"),
            set_prior('normal(1.0, 0.5)', nlpar = "b", lb = 0))
```


```{r}
(formula = bf(median_rating ~ a * saturation ^ b , a ~ 1, b ~ 1 , nl = TRUE))
```


```{r}
# make_stancode(formula, data = df_agg[df_agg$condition == 'unidirectional', ], prior = prior_settings)

```



```{r}
(
  power_law_unidirectional_agg <-
    brm(
      formula,
      data = df_agg[df_agg$condition == 'unidirectional',],
      save_all_pars = T,
      sample_prior = 'yes',
      warmup = 2000,
      iter = 7000,
      cores = 4,
      control = list(adapt_delta = 0.96, max_treedepth = 20),
      prior = prior_settings,
      file = 'study3_power_unidirectional_agg_x'
    )
)
```
